---
title: "Backfiring 2/8/16 update"
author: "mht"
date: "February 8, 2016"
output: html_document
---

Post-cogsci update


## Model motivation

In the backfiring model with the simplest causal, background knoweldge (i.e. intutive theory), people are thought to either have habits or not have habits (e.g. goes for runs), and those give rise to the target behavior (e.g. running on a particular day) with either high or low probability, respectively (called here, **habit strength** and **baserate strength**).

This is the causal knowledge inside the pragmatic listener. The listener's intuitive theory of the speaker is that they can either comment on the event in the affirmative ("John went for a run today.") or negative ("John did not go for a run today."), or stay silent (speaker does not remark on John running). Further, the pragmatic listener believes the speaker to be presupposing that she (the listener) knows the status of that habit (i.e. knows whether or not John usually goes for runs).

The backfiring inference depends on (1) the speaker having the option of not remarking on the situation and (2) the listener believing the speaker to be presupposing that she (the listener) knows the status of the habit.

Given a fixed **habit strength** and **baserate strength**, there is also a (potentially small) predicted effect wherein magnitude of backfiring is expected to be strongest when the *a priori* beliefs are strongest, whereas the mangitude of observational learning is expected to be strongest when the *a priori beliefs are weakest (either, you believe the opposite is likely, or you don't have strong beliefs a priori).



## Experiment 1: Explicit Alternatives

16 items. Each could be instantiated in one of 4 conditions.

1. Communicative / Backfiring
2. Observation
3. Survey / Polar Question
4. Baseline

n = 120. Each participant completed 16 trials, with 4 trials of each of the 4 conditions. 

DV: Forced choice + confidence

Experiment can be viewed [here](http://stanford.edu/~mtessler/backfiring-words/experiments/2afc/forcedchoice-1.html).


```{r data.load, echo=FALSE}
d<-read.csv("~/Documents/research/backfiring-words/data/forcedchoice-2-trials.csv")

d0 <- d %>%
  select(workerid, item, rt, response, condition, confidence) %>%
  mutate(likelyPerson = response=="likely-key",
         condition = factor(condition, 
                       levels=c("prior", "literal", 
                                "pragmatic", "speakermanipulation"),
                       labels=c("baseline","observation","backfire","survey")),
         score = ifelse(likelyPerson, confidence, -confidence))
```


#### Reaction time


```{r RT}
d.rt <- d %>%
  group_by(workerid) %>%
  summarise(rt = mean(rt/1000))

qplot(data=d.rt, x= rt, geom='histogram')+
  xlab("Subject's mean reaction time")

```

RT stats. Mean RT = `r mean(d.rt$rt)`; Median RT = `r median(d.rt$rt)`; SD RT = `r sqrt(var(d.rt$rt))`


#### Distribution of Confidence ratings

```{r confidence}
qplot(data=d0   , x = confidence, geom='histogram', fill=condition)+
  facet_wrap(~condition)
```


#### Overall results

```{r figure, echo=FALSE}
d.fig<-bind_rows(
d0 %>%
  mutate(response = response=="likely-key") %>%
  group_by(condition) %>%
  multi_boot_standard(column="response") %>%
#   rename(fc = mean,
#          fc_lower=ci_lower, 
#          fc_upper=ci_upper) %>%
  mutate(format = "Choice",
         chance =0.5),


d0 %>%
  group_by(condition) %>%
  multi_boot_standard(column="confidence") %>%
#   rename(conf = mean,
#          conf_lower = ci_lower,
#          conf_upper = ci_upper) %>%
  mutate(format = "Confidence",
         chance = 0.5),

d0 %>%
  group_by(condition) %>%
  multi_boot_standard(column="score") %>%
#   rename(conf = mean,
#          conf_lower = ci_lower,
#          conf_upper = ci_upper) %>%
  mutate(format = "Endorsement",
         chance = 0)
) %>%
  mutate(format = factor(format,
                         levels=c("Endorsement","Choice", "Confidence")))

ggplot(d.fig, aes(x=condition, y = mean, fill=condition))+
  geom_bar(stat='identity', width = 0.4, position=position_dodge(0.4))+
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper),
                width=0.3, position=position_dodge(0.4))+
  geom_hline(aes(yintercept = chance), lty=2)+
  theme_paper()+
  #theme(axis.text.y = element_blank())+
#  scale_fill_brewer(palette="Set1")+
  guides(fill=F)+
  #ylim(0,1)+
  ylab("Participant response")+
  facet_wrap(~format)+
  xlab("")+
  coord_flip()

#ggsave("~/Documents/research/backfiring-words/writing/_cogsci2016/figures/expt1-3responses.pdf",
 #      height = 3.5, width = 11)

```



## Planned comparisons


### Analysis 1: Endorsement 1

Fixed effect: Condition

Random effect: Intercept by Subject and Item; Condition by Subject

```{r comparison.score.1}

rs<-lmer(score ~ condition + (1 + condition | workerid) + (1 | item),
          data = d0)

summary(rs)
```

Random effect: Intercept by Subject and Item; Condition by Item

```{r comparison.score.2}
rs.1-lmer(score ~ condition + (1  | workerid) + (1 + condition | item),
          data = d0)

summary(rs.1)
```

WARNING: Correlations between random effect of condition by Item are very high. This makes sense: An item's bias to in a particular condition is very similar to the same item's bias an another condition: There is really only a random effect of intercept for items.

## Analysis 2: Forced choice data

Fixed effect: Condition

Random effect: Intercept by Subject; Condition by Subject

Model fails to converge with more random effects.


```{r comparison.choice}
## maximal model that converges  
rs1<-glmer(likelyPerson~condition + (1 | workerid) + (1 | item),
           family="binomial",
          data = d0)

summary(rs1)

d0 %>%
  mutate(response = response=="likely-key") %>%
  group_by(condition) %>%
  multi_boot_standard(column="response")
```

#### Analysis 3: Confidence data 

```{r comparison.2}
rs2<-lmer(confidence ~ condition + (1 + condition| workerid) + (1 | item),
          data = d0)

summary(rs2)

d0 %>%
  group_by(condition) %>%
  multi_boot_standard(column="confidence")
```

## Explorations


#### Baserates, by item

First, the choice data

```{r explore.baserate}
d.item <- d0 %>%
  group_by(item, condition) %>%
  multi_boot_standard(column="likelyPerson")

qplot(data=d.item %>% filter(condition=="baseline"), x = mean, geom='histogram')+
  xlab("Proportion who choose the *likely* alternative")+
  xlim(0,1)
```

Second, the endorsement scores

```{r explore.baserate2}
d.scr<-d0 %>%
  group_by(item, condition) %>%
  multi_boot_standard(column="score")

qplot(data=d.scr %>% filter(condition=="baseline"), x = mean, geom='histogram')+
  xlab("Average endorsement of *likely* alternative")+
  xlim(-0.5,0.5)
```

Correspondence of proportion choice and endorsement score

```{r explore.baserate3}

d.item.scr<- left_join(d.item %>%
                         rename(prop = mean,
                                prop_lower = ci_lower,
                                prop_upper = ci_upper), 
                       d.scr %>%
                         rename(scr = mean,
                                scr_lower = ci_lower,
                                scr_upper = ci_upper))

ggplot(d.item.scr,aes(x=scr, xmin = scr_lower, xmax= scr_upper,
                      y = prop, ymin =prop_lower, ymax = prop_upper,
                      color = item))+
  geom_point()+
  geom_errorbar()+
  geom_errorbarh()+
  facet_wrap(~condition)+
  xlab("Average score by item")+
  ylab("Proportion *likely* alternative")
```


#### Experimental conditions vs. baseline condition

```{r}
d.item.compare <- left_join(d.item.scr 
                            %>% filter(!(condition =='baseline')),
                            d.item.scr 
                            %>% filter(condition =='baseline') %>%
                              rename(base_prop = prop,
                                     base_prop_lower = prop_lower,
                                     base_prop_upper = prop_upper,
                                     base_scr = scr,
                                     base_scr_lower = scr_lower,
                                     base_scr_uppper = scr_upper) %>%
                              select(-condition))

ggplot(d.item.compare,aes(x=base_scr, xmin = base_scr_lower, xmax= base_scr_uppper,
                      y = scr, ymin =scr_lower, ymax = scr_upper,
                      color = item, group = condition))+
  geom_point()+
  geom_errorbar()+
  geom_errorbarh()+
  geom_smooth()+
  facet_wrap(~condition)+
  xlab("Baseline score")+
  ylab("Condition score")
```


#### Median split of item based on baseline condition

```{r score.medianSplit}

med<-median(d.scr$mean)

d0.medSplit<-left_join(d0, 
  d.scr %>%
    filter(condition=='baseline') %>%
    mutate(medSplit = mean < med) %>%
    mutate(medSplit = factor(medSplit, levels=c("TRUE", "FALSE"),
                             labels=c("weak prior", "strong prior"))) %>%
    select(item, medSplit))

d.scr.med<-d0.medSplit %>%
  group_by(medSplit, condition) %>%
  multi_boot_standard(column="score")

d.scr.med0<-left_join(d.scr.med %>%
  filter(condition != 'baseline') ,
  d.scr.med %>%
    filter(condition == 'baseline') %>%
    rename(prior = mean,
           prior_lower = ci_lower,
           prior_upper = ci_upper) %>%
    select(-condition))

ggplot(d.scr.med, aes(x=condition, y=mean, color=medSplit, group=medSplit))+
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper),
                width=0.1, position=position_dodge(0.4), size=1.5)+
  theme_paper()+
  scale_color_brewer(palette='Set1')+
  ylab("Endorsement score")+
  theme(legend.position="bottom",
        legend.direction="horizontal",
        axis.text.x = element_text(angle=45))

#ggsave("~/Documents/research/backfiring-words/analysis/figures/fc2-score-medSplit.pdf", width=6, height =8)
#ggsave("~/Documents/research/backfiring-words/writing/_cogsci2016/figures//fc2-score-medSplit.pdf", width=5, height =5)
```
