---
title: "Backfire reboot"
output: html_notebook
---

```{r}
project.path <- "~/Documents/research/backfiring-words/"
```

# Prior

```{r}
d.prior <- read.csv( paste(project.path, "data/priors-2-trials.csv", sep = "") )
d.prior.catch <- read.csv (paste (project.path, "data/priors-2-catch_trials.csv", sep = "") )

d.prior <- left_join(d.prior, d.prior.catch %>% select(workerid, pass)) %>%
  filter(pass == 1)
```


```{r}
str(d.prior)
d.p <- d.prior %>% 
  select(workerid, item, 
         ends_with("sample"), ends_with("people"), 
         ends_with("interval"), ends_with("instances"))


d.p.mixtures <- d.p %>%
  select(workerid, item, ends_with("sample"), ends_with("people")) %>%
  gather(key, val, -workerid,-item) %>%
  separate(key, into = c("question", "type")) %>%
  spread(type, val) %>%
  mutate(proportion = people / sample)

# not sure why this throws a warning, no data seems to be lost...
d.p.rates <- d.p %>%
  select(workerid, item, ends_with("interval"), ends_with("instances")) %>%
  gather(key, val, -workerid, -item) %>%
  separate(key, into = c("question", "type")) %>%
  spread(type, val) %>%
  mutate(intervalScale = ifelse(interval == "day", 1,
                         ifelse(interval == "week", 7,
                         ifelse(interval == "month", 30,
                         ifelse(interval == "year", 365,
                         ifelse(interval == "5 years", 365*5, 0))))))

```

Mixture component model

```{r}
mixtureComponentModel <- '
var model = function(){
  var p = uniformDrift({a: 0, b: 1, width: 0.1});
  var obsFn = function(d){
    observe(Binomial({p: p, n: d.sample}), d.people)
  }
  mapData({data:data}, obsFn)
  return {p: p}
}
'
```

Run mixture component of prior model

```{r}
items <- levels(d.p.mixtures$item)
questions <- levels(factor(d.p.mixtures$question))
numSamples <- 20000

rs.fullPost = data.frame()
rs.summary = data.frame()

for (i in items){
  for (q in questions){
    
    d.iq <- d.p.mixtures %>% filter(item == i & question == q)
    
    rs.bda <- webppl(mixtureComponentModel,
                     data = d.iq,
                     data_var = "data",
                     inference_opts = list(method = "MCMC",
                                           samples = numSamples,
                                           burn = numSamples / 2,
                                           verbose = T),
                     model_var = "model",
                     output_format = "samples")
    
    rs.item <- rs.bda %>% mutate(item = i, question = q)
    rs.fullPost <- bind_rows(rs.fullPost, rs.item)
    rs.summary <- bind_rows(rs.summary, rs.item %>%
                              group_by(item, question) %>%
                              summarize(MAP = estimate_mode(p),
                                        cred_upper = hdi_upper(p),
                                        cred_lower = hdi_lower(p)))
    
    print(paste("item = ", i, " ::: question = ", q, sep = ""))
    
  }
}

ggplot(rs.fullPost, aes(x = p))+
  geom_histogram(position = position_dodge())+
  facet_grid(question~item, scales = 'free')

save(rs.summary, 
     file = paste(project.path,
                       "models/results/prior_results/mixtureComponents_summary_mcmc2k.Rdata", sep = ""))
```

Rates model

```{r}
ratesModel <- '
var model = function(){
  var mu = gamma({shape: 0.5, scale: 1});
  var obsFn = function(d){
    observe(Poisson({mu: mu*d.intervalScale}), d.instances)
  }
  mapData({data: data}, obsFn)
  return {mu: mu}
}
'
```

Run rates model

```{r}
items <- levels(d.p.rates$item)
questions <- levels(factor(d.p.rates$question))
numSamples <- 10000

rs.rates.fullPost = data.frame()
rs.rates.summary = data.frame()

for (i in items){
  for (q in questions){
    
    d.iq <- d.p.rates %>% filter(item == i & question == q)
    
    rs.bda <- webppl(ratesModel,
                     data = d.iq,
                     data_var = "data",
                     inference_opts = list(method = "MCMC",
                                           samples = numSamples,
                                           burn = numSamples / 2,
                                           verbose = T),
                     model_var = "model",
                     output_format = "samples")
    
    rs.item <- rs.bda %>% mutate(item = i, question = q)
    rs.rates.fullPost <- bind_rows(rs.rates.fullPost, rs.item)
    rs.summary <- bind_rows(rs.summary, rs.item %>%
                              group_by(item, question) %>%
                              summarize(MAP = estimate_mode(mu),
                                        cred_upper = hdi_upper(mu),
                                        cred_lower = hdi_lower(mu)))
    
    print(paste("item = ", i, " ::: question = ", q, sep = ""))
    
  }
}

save(rs.summary, 
     file = paste(project.path,
                       "models/results/prior_results/mixtureComponents_rates_summary_mcmc2k.Rdata", sep = ""))
```


# Posterior

```{r}
load(paste(project.path,
                       "models/results/prior_results/mixtureComponents_rates_summary_mcmc2k.Rdata", sep = ""))

d.post <- read.csv( paste(project.path, "data/propensity-1-trials.csv", sep = "") )

```

Backfiring RSA model

```{r}
backfireRSA <- '
var limitedPoisson = function(mu){
  var probs = map(function(x){return Math.exp(Poisson({mu: mu}).score(x)) }, [0,1,2,3,4])
  return discrete(probs)
}

var meaning = function(utterance, state){
  return utterance === "yes" ? state > 0 :
          utterance === "no" ? state === 0 :
          true
}

var utterancePrior = function(){
  return uniformDraw(["yes", "mu"])
}

var priorParams = data.priorParams;
var speakerOptimality = data.speakerOptimality[0];

var listener0 = cache(function(priorComponent, utterance, priorParams){
  return Infer({method: "enumerate"}, function(){
    var state = limitedPoisson(priorParams["rates"][priorComponent]);
    var m = meaning(utterance, state);
    condition(m)
    return state
  })
})

var speaker = cache(function(priorComponent, state, speakerOptimality, priorParams){
  return Infer({method: "enumerate"}, function(){
    var utterance = utterancePrior();
    var L0 = listener0(priorComponent, utterance, priorParams);
    factor(speakerOptimality * L0.score(state) )
    return utterance
  })
})

'


# webppl(backfireRSA, 
#        data_var = "data",
#        data = list(
#          priorParams = 
#            list(mixtureProbs = c(0.3, 0.3, 0.4),
#                 rates = c(0.00001, 0.5, 2)),
#          speakerOptimality = 4)
#        )
```


Posterior on "Does it" component

```{r}

l1.v1 <- '
var listener1 = function(utterance, priorParams, speakerOptimality, cost){
  return Infer({method: "enumerate"}, function(){
    var priorComponent = discrete(priorParams.mixtureProbs);
    var state = limitedPoisson(priorParams["rates"][priorComponent]);
    var S1 = speaker(priorComponent, state, speakerOptimality, priorParams);
    observe(S1, utterance)
    var predictive = limitedPoisson(priorParams["rates"][priorComponent])
    return priorComponent
  })
}


Math.exp(listener1("yes", priorParams, speakerOptimality, 1).score(2))
'

rates = list("today." = 1,
               "week." = 7,
               "month." = 30,
               "year." = 365)
getInterval = function(x){
  interval = last(last(strsplit(x, split = " ")))
  return(rates[[interval]])
}

rs.brsa <- data.frame()

for (i in items){
  print(i)

  prior.df <- left_join(
    rs.summary %>% filter(item == i) %>% select(question, item, MAP) %>%
      spread(question, MAP),
    d.post %>% 
      filter(item == i) %>%
      select(item, observationSentence) %>%
      rowwise() %>% 
      mutate(interval = getInterval(as.character(observationSentence))) %>%
      ungroup() %>%
      select(-observationSentence) %>%
      unique()
  )
  
  priorParams <- list(mixtureProbs = c(1 - prior.df$q1, 
                                       prior.df$q1 - prior.df$q2,
                                       prior.df$q2),
                      rates = c(0.00001, 
                                prior.df$interval*prior.df$q4,
                                prior.df$interval*prior.df$q3))
  
  dataToPass <- list(priorParams = priorParams,
                     speakerOptimality = 5)
  
  prob.firstComponent <- webppl(paste(backfireRSA,l1.v1, sep='\n'), data_var = "data", data = dataToPass)
  rs.brsa <- bind_rows(rs.brsa, data.frame(item = i, prior = prior.df$q2, 
                                           posterior = prob.firstComponent))
  
}


ggplot(rs.brsa %>%
         gather(key, val, -item) %>%
         mutate(key = factor(key, levels = c("prior", "posterior"))), 
       aes( x = key, y = val, color = item, group = item))+
  geom_line()+
  facet_wrap(~item)


```


Pedagogy model for comparison

```{r}
pedagogyRSA <- '
var limitedPoisson = function(mu){
  var probs = map(function(x){return Math.exp(Poisson({mu: mu}).score(x)) }, [0,1,2,3,4])
  return discrete(probs)
}

var meaning = function(utterance, state){
  return utterance === "yes" ? state > 0 :
          utterance === "no" ? state === 0 :
          true
}

var utterancePrior = function(){
  return uniformDraw(["yes", "no"])
}

var priorParams = data.priorParams;
var speakerOptimality = data.speakerOptimality[0];

var listener0 = cache(function(utterance){
  return Infer({method: "enumerate"}, function(){
    var priorComponent = discrete(priorParams.mixtureProbs);
    var state = limitedPoisson(priorParams["rates"][priorComponent]);
    var m = meaning(utterance, state);
    condition(m)
    return state
  })
})

var speaker = cache(function(state){
  return Infer({method: "enumerate"}, function(){
    var utterance = utterancePrior();
    var L0 = listener0(utterance);
    factor(speakerOptimality * L0.score(state) )
    return utterance
  })
})

var listener1 = function(utterance){
  return Infer({method: "enumerate"}, function(){
    var priorComponent = discrete(priorParams.mixtureProbs);
    var state = limitedPoisson(priorParams["rates"][priorComponent]);
    var S1 = speaker(state);
    observe(S1, utterance)
    var predictive = limitedPoisson(priorParams["rates"][priorComponent])
    return predictive
  })
}

expectation(listener1("yes"))
'

observationBayes <- '
var limitedPoisson = function(mu){
  var probs = map(function(x){return Math.exp(Poisson({mu: mu}).score(x)) }, [0,1,2,3,4])
  return discrete(probs)
}

var meaning = function(utterance, state){
  return utterance === "yes" ? state > 0 :
          utterance === "no" ? state === 0 :
          true
}

var priorParams = data.priorParams;
var speakerOptimality = data.speakerOptimality[0];

var listener0 = cache(function(utterance){
  return Infer({method: "enumerate"}, function(){
    var priorComponent = discrete(priorParams.mixtureProbs);
    var state = limitedPoisson(priorParams["rates"][priorComponent]);
    var m = meaning(utterance, state);
    condition(m)
    var postPred = limitedPoisson(priorParams["rates"][priorComponent])
    return postPred
  })
})
expectation(listener0("yes"))
'

```

Posterior on rates

```{r}
l1.v2 <- '
var listener1 = function(utterance, priorParams, speakerOptimality, cost){
  return Infer({method: "enumerate"}, function(){
    var priorComponent = discrete(priorParams.mixtureProbs);
    var state = limitedPoisson(priorParams["rates"][priorComponent]);
    var S1 = speaker(priorComponent, state, speakerOptimality, priorParams);
    observe(S1, utterance)
    var predictive = limitedPoisson(priorParams["rates"][priorComponent])
    return predictive
  })
}


expectation(listener1("yes", priorParams, speakerOptimality, 1))
'


rs.brsa <- data.frame()

for (i in items){
  print(i)

  prior.df <- left_join(
    rs.summary %>% filter(item == i) %>%
    select(question, item, MAP) %>%
      spread(question, MAP),
    d.post %>% 
      filter(item == i) %>%
      select(item, observationSentence) %>%
      rowwise() %>% 
      mutate(interval = getInterval(as.character(observationSentence))) %>%
      ungroup() %>%
      select(-observationSentence) %>%
      unique()
  )
  
  priorParams <- list(mixtureProbs = c(1 - prior.df$q1, 
                                       prior.df$q1 - prior.df$q2,
                                       prior.df$q2),
                      rates = c(0.00001, 
                                prior.df$interval*prior.df$q4,
                                prior.df$interval*prior.df$q3))
  
  dataToPass <- list(priorParams = priorParams,
                     speakerOptimality = 5)
  observationExpval <- webppl(observationBayes, data_var = "data", data = dataToPass)

  posteriorExpval <- webppl(paste(backfireRSA,l1.v2, sep='\n'), data_var = "data", data = dataToPass)
  
  pedagogyExpval <- webppl(pedagogyRSA, data_var = "data", data = dataToPass)
  
  
  priorExpval <- sum(priorParams$rates * priorParams$mixtureProbs)
  
  rs.brsa <- bind_rows(rs.brsa, data.frame(item = i, 
                                           prior = priorExpval, 
                                           observation = observationExpval,
                                           pedagogy = pedagogyExpval,
                                           posterior = posteriorExpval,
                                           interval = prior.df$interval,
                                           c1 = priorParams$rates[2],
                                           c2 = priorParams$rates[3]
                                           ))
  
}


ggplot(rs.brsa %>%
         gather(key, val, -item) %>%
         mutate(key = factor(key, levels = c("prior", "posterior"))), 
       aes( x = key, y = val, color = item, group = item))+
  geom_line()+
  facet_wrap(~item, scales = 'free')
```


Run this gives you an "arguments imply different number of rows" error after a certain number of runs. I wonder if this is due to running 3 different webppl calls one after another (some finish before others?), because If I rerun all three webppl calls, they do seem to all get an answer.

Also currently pedagogical and observation give extremely similar / identical answers. 

Is this because of my semantics? That I am operating at the level of instances?

is this because of the state prior being a predictive over counts (as opposed to over rates)?